---
title: "Fundamentals of RMarkdown"
author: "Mike Frank & Chris Hartgerink"
date: "July 31st, 2017"
output:
    revealjs::revealjs_presentation:
        theme: black
        highlight: pygments
        incremental: true
        center: true
        transition: slide
        reveal_options:
            slideNumber: true
        fig_caption: false
bibliography: library.bib
---

```{r, echo = FALSE}
require(statcheck)
```


## Workshop

1. Short intro to Markdown (this!)
2. Create your first RMarkdown file
3. 

## Learning goals

By the end of this workshop you should:

* Know what Markdown is and how the syntax works, 
* See how to integrate code and data with RMarkdown, 
* Understand the different output formats from RMarkdown and how to generate them
* Know about generating APA format files with [`papaja`](https://github.com/crsh/papaja) and bibtex.


## RMarkdown = R + Markdown

1. It combines the magic of R with the simplicity of Markdown documents
1. The purpose: write your manuscripts and automatically insert results!
1. Use cases (amongst others):
  * prevent rounding errors
  * promote computational reproducibility
  * encourage error correction by minimizing manual work

## The Markdown

- Markdown is the simplest document language!
- Open standard, convertable into `.tex`, `.docx`, `.html`, `.pdf`, etc
- You can [learn Markdown in five (!) minutes](https://learnxinyminutes.com/docs/markdown/)
- Despite being simple, very powerful and used on many web platforms

## Basic Markdown: headings

- `# Header 1` creates a heading of the first level
- `## Header 2` creates a heading of the _second_ level
- `### Header 3` ...
- etc

## Basic Markdown: letter styles

- `__bold__` creates __bold__
- `_italic_` creates _italic_
- `***bold and italic!***` creates ***bold and italic!***

## The R

- If you're here, you probably know R
- Programming language for statistics
- Packages to add functionality
- Everything open-source!

## Combining R and Rmarkdown

- This presentation also written in Rmarkdown!
- Can write text, include code
- We can include both code and its outputs, or just one of the two
- Great not only for manuscripts, but also for course notes :-)

## Example code chunk

```{r}
x <- rnorm(100, 50, 20)
y <- rnorm(143, 51, 20)

t.test(x, y)
```

## Example in line

1. However, we can also evaluate code in text, and replace text with output
1. For example, let's round $\pi$ to four decimal places, using the code `round(pi, 4)`
  * `r round(pi, 4)`

## RMarkdown is more powerful

- It's not just about individual results. 
- Ensures data preparation, analysis code is IN the manuscript
- Plus: Don't make all those tables and figures manually! :-)

## Workshop

Work through the handout accompanying these slides, which you can find at 
* bit.ly/rmarkdown-pdf
* bit.ly/rmarkdown-html 
1. [cheat sheet RMarkdown](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)
1. Find this presentation online at [bit.ly/rmarkdown](https://bit.ly/rmarkdown)
