---
title: "RMarkdown for writing reproducible scientific papers"
author: "Mike Frank & Chris Hartgerink"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    number_sections: true
bibliography: mwe.bib
---
  
# Introduction

This is a short tutorial on using RMarkdown to mix prose and code together for creating reproducible scientific documents. It's based on documents that both Chris and Mike wrote about this independently:

[https://elifesciences.org/labs/cad57bcf/composing-reproducible-manuscripts-using-r-markdown]()

and 

[http://babieslearninglanguage.blogspot.com/2015/11/preventing-statistical-reporting-errors.html]()

Now we're together trying to help spread the word. 

## What are we doing? 

Creating documents that can be compiled together with data^[And in a particular computational environment, more on this below.] to produce a full paper manuscript, figures, tables, and all. 

## Why write reproducible papers?

Three reasons. To be right, to be reproducible, and to be efficient. In more depth:

First, to avoid errors. Using an automated method for scraping APA-formatted stats out of PDFs, @nuijten2015 found that over 10% of p-values in published papers were inconsistent with the reported details of the statistical test, and 1.6% were what they called "grossly" inconsistent, e.g. difference between the p-value and the test statistic meant that one implied statistical significance and the other did not. Nearly half of all papers had errors in them. 

Second, to promote computational reproducibility. Computational reproducibility means that other people can take your data and get the same numbers that are in your paper. Even if you don't have errors, it can still be very hard to recover the numbers from published papers because of ambiguities in analysis. Creating a document that literally specifies where all the numbers come from in terms of code that operates over the data removes all this ambiguity.

Third, to create spiffy documents that can be revised easily. This is actually a really big neglected one for us. At least one of us used to tweak tables and figures by hand constantly, leading to a major incentive *never to rerun analyses* because it would mean re-pasting and re-illustratoring all the numbers and figures in a paper. That's a bad thing! It means you have an incentive to be lazy and to avoid redoing your stuff. And you waste tons of time when you do. In contrast, with a reproducible document, you can just rerun with a tweak to the code. You can even specify what you want the figures and tables to look like before you're done with all the data collection (e.g., for purposes of preregistraion or a registered report). 

## Learning goals

By the end of this class you should:

* Know what Markdown is and how the syntax works, 
* See how to integrate code and data in RMarkdown, 
* Understand the different output formats from RMarkdown and how to generate them, and
* Know about generating APA format files with `papaja` and bibtex.

# Markdown syntax

From [http://rmarkdown.rstudio.com/authoring_basics.html](), see also [this cheat sheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf). 

The idea of markdown is that it's really really easy to read and write:

* It's easy to get `*italic*` or `**bold**`. 
* You can get headings using `# heading1` for first level, `## heading2` for second-level, and `### heading3` for third level. 
* Lists are delimited with `*` for each entry.
* You can write links by writing `[here's my link](http://foo.com)`.

And then you can insert code by using the code chunk environment `\`\`\`\`, and its outputs will get mixed in. 

# Code and prose, mixed together

First load packages, I always put these together at the top. 

```{r}
library(knitr)
library(ggplot2)
library(broom)
library(devtools)
```

Now, here are some of the startup options I often use. Caching can be very helpful for large files, but can also cause problems when there are external dependencies that change. 

```{r}
opts_chunk$set(fig.width=8, fig.height=5, 
                      echo=TRUE, warning=FALSE, message=FALSE, cache=TRUE)
```

And you can use various local and glbal chunk options like `echo=FALSE` to suppress showing the code (better for papers). 

## Graphs

It's really easy to include graphs, like this one.

```{r}
qplot(hp, mpg, col = factor(cyl), data = mtcars)
```

## Statistics

It's also really easy to include statistical tests of various types. 

For this I really like the `broom` package, which formats the outputs of various tests really nicely. Paired with knitr's `kable` you can make very simple tables. 

```{r}
mod <- lm(mpg ~ hp + cyl, data = mtcars)
kable(tidy(mod), digits = 3)
```

Of course, cleaning these up can take some work. For example, we'd need to rename a bunch of fields to make this table have the labels we wanted (e.g., to turn `hp` into `Horsepower`). 

I also do a lot of APA-formatted statistics. We can compute them first, and then print them inline.

```{r}
ts <- with(mtcars,t.test(hp[cyl==4], hp[cyl==6]))
```

There's a statistically-significant difference in horsepower for 4- and 6-cylinder cars  ($t(`r  round(ts$parameter,2)`) = `r  round(ts$statistic,2)`$, $p = `r round(ts$p.value,3)`$). To insert these stats inline I wrote e.g. `round(ts$parameter, 2)` inside an inline code block.

Note that rounding can get you in trouble here, because it's very easy to have an output of $p = 0$ when in fact $p$ can never be exactly equal to 0. 

# Output Formats


# Writing APA-format papers

## Using the `papaja` package



## Bibiographic management

It's also possible to include references using `bibtex`, by using `@ref` syntax. I like [bibdesk](http://bibdesk.sourceforge.net/) as a simple reference manager that integrates with google scholar, but many other options are possible. 

With a bibtex file included, you can refer to papers like @nuijten2016 in text, or cite them parenthetically [e.g., @nuijten2016].

So in conclusion, and as described by @xie2013, `knitr` is really amazing! 

## Computational reproducibility concerns

`packrat` for package versioning

or include `session_info` from `devtools` at least. 

```{r}
devtools::session_info()
```

could even use `docker`, but that's outside the scope of the course!

# Bibliography
